---
title: "Collision with holiday and weather"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library

```{r}
library(dplyr)
#install.packages("dlookr")
#install.packages("Rttf2pt1")
#install.packages("inum")
library(dlookr)
library(tidyverse)
#install.packages("flextable")
# install.packages("gdtools")
#library(flextable)
library(tidyr)
```

## Data Preparation

### Capstone time series with holiday and weather analysis

```{r}
data = read.csv("collision_w_holidays_n_weather.csv")
```

### Data quality
```{r}
data$CRASH.DATE = as.Date(data$CRASH.DATE)
data_12_to_23 = subset(data, data$CRASH.DATE >= as.Date("2012-01-01") & data$CRASH.DATE <= as.Date("2023-12-31"))

summary(data_12_to_23)
```
```{r}
diagnose(data_12_to_23) %>%
  select(-unique_count, -unique_rate)%>%
  filter(missing_count != 0)%>%
  arrange(desc(missing_percent))%>% 
  knitr::kable(align = 'c', format = "markdown")
```
```{r}
data_12_to_23 = subset(data_12_to_23, select = -c(Peak_gust_wind_speed, sunshine_duration))

```


#### This one contains the columns that have missing values

```{r}
diagnose(data_12_to_23) %>%
  select(-unique_count, -unique_rate)%>%
  filter(missing_count != 0)%>%
  arrange(desc(missing_percent))%>% 
  knitr::kable(align = 'c', format = "markdown")
```

#### after delete those unnecessary row, here is the whole data quality report the this data set

```{r}
diagnose(data_12_to_23)%>% 
  knitr::kable(align = 'c', format = "markdown")
```

#### diagnose category data and numeric data seperatly

```{r}
diagnose_category(data_12_to_23)%>% 
  knitr::kable(align = 'c', format = "markdown")
```
```{r}
diagnose_numeric(data_12_to_23)%>% 
  knitr::kable(align = 'c', format = "markdown")
```

### Add a new column: weekday/weekend

```{r}
data_12_to_23 <- data_12_to_23 %>%
  mutate(day_type = ifelse(weekdays(CRASH.DATE) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))

```

## Weather

### snowfall amount vs people injured

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_people_injured = mean(NUMBER.OF.PERSONS.INJURED))%>%
  replace_na(list(average_people_injured = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_people_injured)) +
  geom_bar(stat = "identity", width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of People Injured")

summary_table
```

### snowfall amount vs people killed

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_people_killed = mean(NUMBER.OF.PERSONS.KILLED))%>%
  replace_na(list(average_people_killed = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_people_killed)) +
  geom_bar(stat = "identity", width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of People Killed")

summary_table
```

### snowfall amount vs pedestrians injured

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_pedestrians_injured = mean(NUMBER.OF.PEDESTRIANS.INJURED))%>%
  replace_na(list(average_pedestrians_injured = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_pedestrians_injured)) +
  geom_bar(stat = "identity", width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of Pedestrians Injured")

summary_table
```

### snowfall amount vs pedestrians killed

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_pedestrians_killed = mean(NUMBER.OF.PEDESTRIANS.KILLED ))%>%
  replace_na(list(average_pedestrians_killed = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_pedestrians_killed)) +
  geom_bar(stat = "identity", width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of Pedestrians Killed")

summary_table
```
### snowfall amount vs cyclist injured

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_cyclist_injured = mean(NUMBER.OF.CYCLIST.INJURED))%>%
  replace_na(list(average_cyclist_injured = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_cyclist_injured)) +
  geom_bar(stat = "identity",width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of Cyclist Injured")

summary_table
```
### snowfall amount vs cyclist killed

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_cyclist_killed = mean(NUMBER.OF.CYCLIST.KILLED  ))%>%
  replace_na(list(average_cyclist_killed = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_cyclist_killed)) +
  geom_bar(stat = "identity", width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of Cyclist Killed")

summary_table
```

### snowfall amount vs motorist injured

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_motorist_injured = mean(NUMBER.OF.MOTORIST.INJURED ))%>%
  replace_na(list(average_motorist_injured = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_motorist_injured)) +
  geom_bar(stat = "identity", width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of Motorist Injured")

summary_table
```

### snowfall amount vs motorist killed

```{r}
library(ggplot2)

summary_table = data_12_to_23 %>%
  group_by(snowfall_amount) %>%
  summarize(average_motorist_killed = mean(NUMBER.OF.MOTORIST.KILLED   ))%>%
  replace_na(list(average_motorist_killed = 0))

ggplot(summary_table, aes(x = snowfall_amount, y = average_motorist_killed)) +
  geom_bar(stat = "identity", width = 3,color="steelblue",fill="steelblue") +
  labs(x = "Snowfall Amount", y = "Number of Motorist Killed")

summary_table
```
## Date type

### Weekday weekend and holiday VS people injured

```{r}
data_12_to_23 <- data_12_to_23 %>%
  mutate(Holiday = na_if(Holiday, ""))

data_12_to_23 <- data_12_to_23 %>%
  mutate(day_type_holiday = ifelse(!is.na(Holiday), "Holiday", day_type)) 

weekday_weekend_holiday <- data_12_to_23 %>%
  mutate(day_type1 = ifelse(!is.na(Holiday), "Holiday", day_type)) %>%
  group_by(day_type1) %>%
  summarize(count = n())

weekday_weekend_holiday_injury <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_injured = sum(NUMBER.OF.PERSONS.INJURED, na.rm = TRUE),
            average_injured = mean(NUMBER.OF.PERSONS.INJURED, na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_injury, aes(x = day_type_holiday, y = total_injured, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total Injured", title = "Total Injured by Day Type") +
  geom_text(aes(label = total_injured), vjust = -0.5, size = 3) +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_injured*1000000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_injured*1000000, label = round(average_injured, 3)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./1000000, name = "Average Injured"))+scale_fill_brewer(palette = "Set2")

```

### Weekday weekend and holiday VS people killed

```{r}
weekday_weekend_holiday_killed <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_killed = sum(NUMBER.OF.PERSONS.KILLED, na.rm = TRUE),
            average_killed = mean(NUMBER.OF.PERSONS.KILLED, na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_killed, aes(x = day_type_holiday, y = total_killed, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total Killed", title = "Total Killed by Day Type") +
  geom_text(aes(label = total_killed), vjust = -0.5, size = 3) +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_killed*1000000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_killed*1000000, label = round(average_killed, 5)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./1000000, name = "Average Killed"))+scale_fill_brewer(palette = "Set2")
```

### Weekday weekend and holiday VS pedestrians injured

```{r}
weekday_weekend_holiday_injured <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_injured = sum(NUMBER.OF.PEDESTRIANS.INJURED, na.rm = TRUE),
            average_injured = mean(NUMBER.OF.PEDESTRIANS.INJURED, na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_injured, aes(x = day_type_holiday, y = total_injured, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total injured", title = "Total injured by Day Type") +
  geom_text(aes(label = total_injured), vjust = -0.5, size = 3) +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_injured*1000000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_injured*1000000, label = round(average_injured, 3)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./1000000, name = "Average Injured"))
```

### Weekday weekend and holiday VS pedestrians killed

```{r}
weekday_weekend_holiday_killed <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_killed = sum(NUMBER.OF.PEDESTRIANS.KILLED, na.rm = TRUE),
            average_killed = mean(NUMBER.OF.PEDESTRIANS.KILLED, na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_killed, aes(x = day_type_holiday, y = total_killed, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total Killed", title = "Total Killed by Day Type") +
  geom_text(aes(label = total_killed), vjust = -0.5, size = 3) +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_killed*1000000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_killed*1000000, label = round(average_killed, 5)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./1000000, name = "Average Killed"))
```

### Weekday weekend and holiday VS Cyclist injured

```{r}
weekday_weekend_holiday_injured <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_injured = sum(NUMBER.OF.CYCLIST.INJURED , na.rm = TRUE),
            average_injured = mean(NUMBER.OF.CYCLIST.INJURED , na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_injured, aes(x = day_type_holiday, y = total_injured, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total injured", title = "Total injured by Day Type") +
  geom_text(aes(label = total_injured), vjust = -0.5, size = 3) +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_injured*1000000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_injured*1000000, label = round(average_injured, 4)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./1000000, name = "Average Injured"))
```

### Weekday weekend and holiday VS cyclist killed

```{r}
weekday_weekend_holiday_killed <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_killed = sum(NUMBER.OF.CYCLIST.KILLED, na.rm = TRUE),
            average_killed = mean(NUMBER.OF.CYCLIST.KILLED, na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_killed, aes(x = day_type_holiday, y = total_killed, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total Killed", title = "Total Killed by Day Type") +
  geom_text(aes(label = total_killed), vjust = -0.5, size = 3) +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_killed*700000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_killed*700000, label = round(average_killed, 6)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./700000, name = "Average Killed"))
```

### Weekday weekend and holiday VS Motorist injured

```{r}
weekday_weekend_holiday_injured <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_injured = sum(NUMBER.OF.MOTORIST.INJURED , na.rm = TRUE),
            average_injured = mean(NUMBER.OF.MOTORIST.INJURED , na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_injured, aes(x = day_type_holiday, y = total_injured, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total injured", title = "Total injured by Day Type") +
  geom_text(aes(label = total_injured), vjust = -0.5, size = 3) +
  geom_line(aes(x = day_type_holiday, y = average_injured*1000000, group = 1), color = "blue") +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_injured*1000000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_injured*1000000, label = round(average_injured, 4)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./1000000, name = "Average Injured"))
```

### Weekday weekend and holiday VS motorist killed

```{r}
weekday_weekend_holiday_killed <- data_12_to_23 %>%
  group_by(day_type_holiday) %>%
  summarize(total_killed = sum(NUMBER.OF.MOTORIST.KILLED, na.rm = TRUE),
            average_killed = mean(NUMBER.OF.MOTORIST.KILLED, na.rm = TRUE))
```

```{r}
ggplot(weekday_weekend_holiday_killed, aes(x = day_type_holiday, y = total_killed, fill = day_type_holiday)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(x = "Day Type", y = "Total Killed", title = "Total Killed by Day Type") +
  geom_text(aes(label = total_killed), vjust = -0.5, size = 3) +
  theme_minimal() +
  geom_line(aes(x = day_type_holiday, y = average_killed*700000, group = 1), color = "blue") +
  geom_text(aes(x = day_type_holiday, y = average_killed*700000, label = round(average_killed, 6)), vjust = -1, size = 3, color = "black") +
  scale_y_continuous(sec.axis = sec_axis(~./700000, name = "Average Killed"))
```

